---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
data <- read.csv("activity.csv")
head(data)
```


## What is mean total number of steps taken per day?

### 1. Calculate the total number of steps taken per day

```{r message=FALSE, warning=FALSE}
library(dplyr)
data_steps_taken_perday <- data %>%
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        summarize(PerDaySteps = sum(steps), steps = steps)

data_steps_taken_perday        
```

### 2. Histogram for steps taken per day

```{r}
library(ggplot2)
ggplot(data_steps_taken_perday, aes(x = PerDaySteps)) + 
  geom_histogram(binwidth = 1000) + 
  xlab("Per Day Steps") +
  ggtitle("Steps Taken Per Day") +
  theme(plot.title = element_text(hjust = 0.5))

```

### 3. Calculate and report the mean and median of the total number of steps taken per day

```{r message=FALSE, warning=FALSE}
data %>%
        filter(!is.na(steps)) %>%
        group_by(date) %>%
        summarize(mean = mean(steps, na.rm = TRUE), 
                  median = median(steps, na.rm = TRUE), 
                  total_steps = sum(steps, na.rm = TRUE))
```


## What is the average daily activity pattern?

### 1. Make a time series plot (i.e. type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r message=FALSE, warning=FALSE}
data_daily_activity_pattern <- data %>%
        filter(!is.na(steps)) %>%
        group_by(interval) %>%
        summarize(mean = mean(steps, na.rm = TRUE))

ggplot(data_daily_activity_pattern, aes(x = interval, y = mean)) + geom_line()+
  xlab("5-minute interval") +
  ylab("Avg. Steps") +
  ggtitle("5-minute interval Vs Avg. Number of Steps Taken") +
  theme(plot.title = element_text(hjust = 0.5))
  
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

``` {r message=FALSE, warning=FALSE}
data_daily_activity_pattern %>%
  filter(mean == max(mean))
```

## Imputing missing values

### 1. Report of missing values

``` {r message=FALSE, warning=FALSE}
data %>%
  summarize(total = n(), na = sum(is.na(steps)), na_percent = (na / total) * 100)
```

### 2. impute missing values
We will use mean to impute the missing values.

```{r message=FALSE, warning=FALSE}
data_after_impute <- data %>%
  mutate(steps = replace(steps, is.na(steps), mean(steps, na.rm = TRUE)))

head(data_after_impute)

```

Histogram with the new data set after imputing missing values.

```{r message=FALSE, warning=FALSE}
data_after_impute_hist <- data_after_impute %>%
        group_by(date) %>%
        summarize(PerDaySteps = sum(steps), steps = steps)

ggplot(data_after_impute_hist, aes(x = PerDaySteps)) + geom_histogram(binwidth = 1000) + 
  xlab("Per Day Steps") + 
  ggtitle("Steps Taken Per Day After Imputing Missing Value") +
  theme(plot.title = element_text(hjust = 0.5))
```

Observation: After impute the missing values we can see from the histogram that total count of steps per day has increased significantly.

Report mean, median & total number of steps taken per day

```{r message=FALSE, warning=FALSE}
data_after_impute %>%
        group_by(date) %>%
        summarize(mean = mean(steps, na.rm = TRUE), 
                  median = median(steps, na.rm = TRUE), 
                  total_steps = sum(steps))
```

Observation: After impute the missing values mean, median & total steps increase.

## Are there differences in activity patterns between weekdays and weekends?

### 1. Apply weekend and weekday factor

```{r message=FALSE, warning=FALSE}
library(lubridate)
data_with_factor <- data_after_impute %>%
  mutate(weekday_weekend = 
           factor(ifelse(wday(as_date(date)) %in% c(1,7), "weekend", "weekday")))
head(data_with_factor)
```

### 2. Panel plot by factor

```{r message=FALSE, warning=FALSE}
data_factorized <- data_with_factor %>%
  group_by(interval) %>%
  summarize(avg_step_perday = mean(steps), weekday_weekend)

plot <- ggplot(data_factorized, aes(x = interval, y = avg_step_perday)) + 
  geom_line() + 
  facet_grid(row = vars(weekday_weekend)) +
  xlab("5-minute interval") + 
  ylab("Avg. Steps") +
  ggtitle("5-minute interval Vs Steps Taken Per Day") +
  theme(plot.title = element_text(hjust = 0.5))

png(file = "figure/weekday_weekend_comparison.png")
print(plot)
dev.off()
    
```

